---
title: "Diamond Data Visualization with R"
---
```{r}

library(ggplot2)
data(diamonds)
names(diamonds)
```

```{r}
summary(diamonds)
```
### Representing the best Color
```{r}
qplot(data = diamonds, x = color, color = I('black'), fill = I('#F79420'),
      xlab = 'Color', ylab = 'Number of Diamonds')
```
### Creating a histogram of the price of all the diamonds in the diamond data set.
```{r}
qplot(data = diamonds, x = price, color = I('black'), fill = I('#F79420'), 
      xlab = 'Price', ylab = 'Number of Diamonds')  +
  ggtitle('Diamond Price') + 
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
summary(diamonds$price)
```
### How many diamonds cost less than $500?
```{r}
nrow(subset(diamonds, price < 500))
```
### How many diamonds cost less than $250?
```{r}
nrow(subset(diamonds, price < 250))
```
### How many diamonds cost $15000 or more?
```{r}
nrow(subset(diamonds, price <= 15000))
```
### Changing the transparency of the points and making the x-axis every 2 units
```{r}
 ggplot(diamonds) +
  geom_histogram( aes(x = price), color = I('black'), fill = I('#F79420'), 
      xlab = 'Price', ylab = 'Number of Diamonds')  +
  ggtitle('Diamond Price') + 
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0,19000))
```

### Breaking down the main price histogram to five histograms in separate panels.
```{r}
 ggplot(diamonds) +
  geom_histogram( aes(x = price), color = I('black'), fill = I('#F79420'), 
      xlab = 'Price', ylab = 'Number of Diamonds')  +
  ggtitle('Diamond Price') + 
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0,19000)) +
  facet_grid(. ~ cut)
```

### Investigate the price of diamonds using box plots

```{r}
ggplot(diamonds) +
  geom_boxplot(aes(x = color,  y = price, fill = color)) +
   ggtitle("Diamonds Price by Color.") +
   theme(plot.title = element_text(hjust = 0.5))
```

### what is the price range for the middle 50% of diamonds with color D?
### what is the price range for the middle 50% of diamonds with color J?
### what is the IQR for the diamonds with best color?
### what is the IQR for the diamonds with worst color?

```{r}
summary(diamonds$price[diamonds$color == 'D'])
summary(diamonds$price[diamonds$color == 'J'])
IQR(diamonds$price[diamonds$color == "D"])
IQR(diamonds$price[diamonds$color == "J"])
```

### Investigating the price per carat of diamonds across the different colors of diamonds using boxplots.

```{r}
ggplot(diamonds) + 
  geom_boxplot(aes(x = color,  y = price/carat, fill = color)) +
   ggtitle("Diamonds Price by Color.")

```
### investigate the weight of the diamonds using frequency polygon.
### what carat size has a count greater than 2000?

```{r}

ggplot(diamonds) + geom_freqpoly(aes(x = carat), binwidth = 0.5) +
scale_x_continuous(limits = c(0,6), breaks = seq(0,6,0.1))


```


