---
title: "Diamond Data Visualization with R"
output: pdf_document
---
```{r}

library(ggplot2)
data(diamonds)
names(diamonds)
```

```{r}
summary(diamonds)
```
### Representing the best Color
```{r}
qplot(data = diamonds, x = color, color = I('black'), fill = I('#F79420'),
      xlab = 'Color', ylab = 'Number of Diamonds')
```
### Creating a histogram of the price of all the diamonds in the diamond data set.
```{r}
qplot(data = diamonds, x = price, color = I('black'), fill = I('#F79420'), 
      xlab = 'Price', ylab = 'Number of Diamonds')  +
  ggtitle('Diamond Price') + 
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
summary(diamonds$price)
```
### How many diamonds cost less than $500?
```{r}
nrow(subset(diamonds, price < 500))
```
### How many diamonds cost less than $250?
```{r}
nrow(subset(diamonds, price < 250))
```
### How many diamonds cost $15000 or more?
```{r}
nrow(subset(diamonds, price <= 15000))
```
### Changing the transparency of the points and making the x-axis every 2 units
```{r}
 ggplot(diamonds) +
  geom_histogram( aes(x = price), color = I('black'), fill = I('#F79420'), 
      xlab = 'Price', ylab = 'Number of Diamonds')  +
  ggtitle('Diamond Price') + 
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0,19000))
```

### Breaking down the main price histogram to five histograms in separate panels.
```{r}
 ggplot(diamonds) +
  geom_histogram( aes(x = price), color = I('black'), fill = I('#F79420'), 
      xlab = 'Price', ylab = 'Number of Diamonds')  +
  ggtitle('Diamond Price') + 
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_cartesian(xlim = c(0,19000)) +
  facet_grid(. ~ cut)
```

### Investigate the price of diamonds using box plots

```{r}
ggplot(diamonds) +
  geom_boxplot(aes(x = color,  y = price, fill = color)) +
   ggtitle("Diamonds Price by Color.") +
   theme(plot.title = element_text(hjust = 0.5))
```

### what is the price range for the middle 50% of diamonds with color D?
### what is the price range for the middle 50% of diamonds with color J?
### what is the IQR for the diamonds with best color?
### what is the IQR for the diamonds with worst color?

```{r}
summary(diamonds$price[diamonds$color == 'D'])
summary(diamonds$price[diamonds$color == 'J'])
IQR(diamonds$price[diamonds$color == "D"])
IQR(diamonds$price[diamonds$color == "J"])
```

### Investigating the price per carat of diamonds across the different colors of diamonds using boxplots.

```{r}
ggplot(diamonds) + 
  geom_boxplot(aes(x = color,  y = price/carat, fill = color)) +
   ggtitle("Diamonds Price by Color.")

```
### investigate the weight of the diamonds using frequency polygon.
### what carat size has a count greater than 2000?

```{r}

ggplot(diamonds) + geom_freqpoly(aes(x = carat), binwidth = 0.5) +
scale_x_continuous(limits = c(0,6), breaks = seq(0,6,0.1))


```

### Create a scatterplot of price vs x
```{r}
ggplot(data = diamonds) +
       xlab('Length in mm') +
       ylab('Price in USD') + 
  geom_point(aes(x = x, y = price),color = 'blue', alpha = 1/10)
```

### Correlation between price and x, between price and y, between price and z
```{r}
cor.test(diamonds$price,diamonds$x)
```

```{r}
cor.test(diamonds$price,diamonds$y) 
```

```{r}
cor.test(diamonds$price,diamonds$z)
```

### Create a scatterplot of price vs depth

```{r}
ggplot(data = diamonds) +
       xlab('Depth in mm') +
       ylab('Price in USD') + 
  geom_point(aes(x = depth, y = price),alpha=1/100, color = 'blue') +
  scale_x_continuous(limits = c(50, 75),
                     breaks = seq(50,75, 2))
```
### Correlation between price and depth
```{r}
cor.test(diamonds$price,diamonds$depth)
```

### Creating a scatterplot of price vs carat and omit the top 1% of price and carat values.

```{r}
ggplot( data = diamonds) +
       xlab('Carat') +
       ylab('Price in USD') + 
  geom_point(aes(x = carat, y = price),color = 'blue') +
  xlim(0, quantile(diamonds$carat, probs = 0.99)) +   # 99% percentile on x variable
  ylim(0, quantile(diamonds$price, probs = 0.99))
```

### Create a scatterplot of price vs volume (xyz)

```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot( data = diamonds) +
       xlab('Volume (mm3)') +
       ylab('Price in USD') + 
  geom_point(aes(x = volume, y = price),color = 'blue') 
```

### number of diamonds with volume=0

```{r}
dim(diamonds[diamonds$volume == 0,])
```

### Correlation between price vs volume (xyz)
### Exclude diamonds with volume=0 and volume >= 800

```{r}
temp_df <- diamonds[ diamonds$volume > 0 & diamonds$volume <= 800 , ]

cor.test(temp_df$price,temp_df$volume)
```

### Create a scatterplot of price vs volume (xyz)
```{r}
temp_df <- diamonds[ diamonds$volume > 0 & diamonds$volume <= 800 , ]

ggplot( data = temp_df) +
       xlab('Volume (mm3)') +
       ylab('Price in USD') + 
  geom_point(aes(x = volume, y = price),color = 'blue',alpha=1/100) +
  geom_smooth(aes(x = volume, y = price))
```
### Create a new data frame containing info on diamonds by clarity

```{r}
library(dplyr)

clarity_groups  <- group_by(diamonds,clarity)     # first groups data by clarity

diamondsByClarity <- summarise(clarity_groups,                 # then summarizes
          mean_price = mean(price),                    # and creates new variables
          median_price = median(price),
          min_price = min(price),
          max_price = max(price),
          n = n()                                      # number of items in each group
          )

diamondsByClarity
```
### Create two bar plots : mean_price vs. clarity, mean_price vs. color

```{r}
library(dplyr)
library(ggplot2)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))
p1 <- ggplot( data = diamonds_mp_by_clarity) +
  geom_bar(aes(x = clarity, y = mean_price),stat="identity",color = 'black', fill = 'orange')

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
p2 <- ggplot( data = diamonds_mp_by_color) +
  geom_bar(aes(x = color, y = mean_price),stat="identity",color = 'black',fill = 'blue')

library(gridExtra)
grid.arrange(p1,p2,ncol=1)
```
```{r}
library(ggplot2)
data(diamonds)
summary(diamonds)
```

### Create a histogram of diamond prices

```{r}
ggplot(data = diamonds) + 
  geom_histogram(aes(fill = cut, x=price) ) +
  facet_wrap(~color) +
  ggtitle('Diamond prices, by color and cut')
```

### Create a scatterplot of diamond prices

```{r}
ggplot( data = diamonds) + 
  geom_point(aes(x=table, y=price,color = cut),alpha=1/3) +
  scale_x_continuous(limits = c(50, 70),
                     breaks = seq(50, 70, 5)) + 
  ggtitle('Diamond prices as a function of table, by cut')
```

### Create a scatterplot of diamond prices

```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(data = diamonds) + 
  geom_point(alpha=1/5, aes(color = clarity, x=volume, y=price)) +
  xlim(0, quantile(diamonds$volume, probs = 0.99)) +   
  scale_y_log10() + 
  ggtitle('Diamond prices as a function of volume, by clarity \n (omitting top 1% of diamond volumes)')
```

